---
title: "11. Social Network Analysis"
output: html_notebook
---

##노드와 연결선으로 표현 가능한 관계형 데이터에 대해 사용가능한 분석기법

###Social Network MAP
####노드와 엣지로 구성.
    -> 크기, 두께 : 중요성, 관계의 정도.
    -> 엣지: 방향성(전화를 거는 방향), 무방향성(Bipartite->Unipartite).

###SNA Graphing method in R.
	  1. plot( graph.edgelist( mat[start, end] )
	  2. gplot( relation matrix, )


###사례1. 전화
####String - Factorize - Integer
####graph object
####edgelist(X)
#### plot(tele.w,...)
```{r}
options(warn = -1)
library(igraph)
getwd()
tele <- read.csv('./머신러닝_독자제공용자료모음/Practice/tele.csv', header = FALSE)
head(tele)
levels_all <- union(levels(tele$V1) , levels(tele$V2))
tele$from <- factor(tele$V1 , levels = levels_all) 
tele$to <- factor(tele$V2 , levels = levels_all)
tele$from2 <- as.integer(tele$from)
tele$to2 <- as.integer(tele$to) 
tele_mat <- cbind(from = tele$from2 , to = tele$to2 , cnt = tele$V3) 
tele_mat
tele.w <- graph.edgelist(tele_mat[,1:2])
E(tele.w)$weight <- tele_mat[,3]
str(tele.w)

tele.diag <- rep(0,16) + 5
tele.name <- levels_all
tele

#Does not work.

# plot(tele.w,
#      vertex.size=10,vertex.shape="circle",vertex.size=tele.diag,
#      vertex.label=tele.name,vertex.label.font=1,
#      vertex.label.cex=1+sqrt(tele.diag)/15,
#      edge.width=2+E(tele.w)$weight/2, edge.arrow.width= E(tele.w)$weight/50)

plot(tele.w,
     vertex.size=10,vertex.shape="circle",vertex.size=tele.diag,
     vertex.label=tele.name,vertex.label.font=1,
     vertex.label.cex=1+sqrt(tele.diag)/15,
     edge.width=2+tele_mat[,3]/2, edge.arrow.width= tele_mat[,3]/50)


```
#메일 관계
## String - Factorize - Integer
## Graph object(edgelist matrix) : email.w
## plot(email.w,...)
```{r}
email <- read.csv("./머신러닝_독자제공용자료모음/Practice/email.csv",header=F)
levels_all <- union(levels(email$V1) , levels(email$V2))
email$from <- factor(email$V1 , levels = levels_all) 
email$to <- factor(email$V2 , levels = levels_all) 
email$from2 <- as.integer(email$from) 
email$to2 <- as.integer(email$to) 
email_mat <- cbind(from = email$from2 , to = email$to2 , cnt = email$V3) 
email
email_mat
email.w <- graph.edgelist(email_mat[,1:2])
E(email.w)$weight <- email_mat[,3]
str(email.w)

email.diag <- rep(0,16) + 5
email.name <- levels_all

plot(email.w,  
     vertex.size=10,vertex.shape="circle",vertex.size=email.diag,
     vertex.label=email.name,vertex.label.font=1,
     vertex.label.cex=1+sqrt(email.diag)/15,
     edge.width=2+email_mat[,3]/2, edge.arrow.width= email_mat[,3]/50)
```
#Bipartite Network
##도서구입관계
##Bipartite Network.
## x, y축의 모든 요소를 하나의 축으로 가지는 Matrix.

```{r}

#BOOK
book <- read.csv("./머신러닝_독자제공용자료모음/Practice/book.csv", header=T, stringsAsFactors = F)
book
book[is.na(book)] = 0
book
rownames(book) <- book[,1]
book <- book[-1]
book

book_mat <- as.matrix(book)
n <- nrow(book_mat)
m <- ncol(book_mat)
book_mat2 <- rbind(cbind(matrix(0,n,n),book_mat),cbind(t(book_mat),matrix(0,m,m)))

```

```{r}

library(sna)
vertex.col <- c(rep("blue",n),rep("green",m))
vertex.cex <- c(rep(2,n),rep(2,m))

gplot(book_mat2, mode="circle", displaylabels=T, boxed.labels=F, 
      vertex.col=vertex.col, vertex.cex=vertex.cex, 
      label.col=vertex.col, label.cex=1.2, usearrows=F)

```
## Bipartite -> Unipartite(구입자간 관계도)
## Squares of Density
### ex) 현주{books} * {books} 현주
###Diag : 자신이 산 책수의 제곱합
## 반대로 책들간 관계도 가능
### ex) Hadoop {구입자} * {구입자} Hadoop
## Density (in == out)

```{r}
#book_mat
#
book_mat3 <- book_mat %*% t(book_mat)
#book_mat3

diag(book_mat3) <- rep(0 , dim(book_mat3)[1])
book_mat3[book_mat3 > 0] <- 1
#book_mat3

gplot(book_mat3, mode="circle", displaylabels=T, boxed.labels=F, 
      vertex.col="blue", vertex.cex=2, 
      label.col="green", label.cex=1.2, usearrows=F)

#Network matrix connecting from M(i,j) : connectivity of (i-j th node element)
rownames(book_mat3)
degree(book_mat3) #Undirected, Without in/out degree
degree(book_mat3 , cmode ="indegree")
degree(book_mat3 , cmode ="outdegree")

#Density and Reacheability of each node
gden(book_mat3) #Find Graph density
reachability(book_mat3)

```
#네트워크 용어
## Component
    -> 각각의 노드들이 닿을 수 있는 노드리스트의 세트?

```{r}
#무작위 그래프샘플
set.seed(123)
yang_1 <- rgraph(10 , tprob = 1)
gplot(yang_1 , displaylabels = T , boxed.labels = F , vertex.cex = 2) 
yang_03 <- rgraph(10, tprob = 0.2)

gplot(yang_03, displaylabels = T, boxed.labels = F, vertex.cex = 2)
components(yang_03)
component.dist(yang_03)

#Hamming Distance of Sub-Graphs (?)
sedist(yang_03 , method = "hamming")
cluster_03 <- equiv.clust(yang_03 , method = "hamming" , cluster.method = "complete")
plot(cluster_03)

bplot <- blockmodel(yang_03 , cluster_03 , h = 3)
plot(bplot)
bplot

yang_03
geodist(yang_03)$gdist

closeness(book_mat3)
rownames(book_mat3)[which.max(closeness(book_mat3))]

```

```{r}
#변환
range(tele_mat[,2])
tele_mat2 <- matrix(0 , 12 , 12)

#Convert into relation matrix
#  == # tele_mat2[tele_mat[,1:2]] <- 1
for(i in 1:16) {
  tele_mat2[as.integer(tele_mat[i,][1]) , as.integer(tele_mat[i,][2])] = as.integer(tele_mat[i,][3])  
}

#Naming with factor levels ? random leveling?
rownames(tele_mat2) <- union(levels(tele$V1) , levels(tele$V2))
colnames(tele_mat2) <- union(levels(tele$V1) , levels(tele$V2))
#tele_mat2

gplot(tele_mat2, mode="circle", displaylabels=T, boxed.labels=F, 
      vertex.col="blue", vertex.cex=2, edge.lwd = tele_mat2 , 
      label.col="green", label.cex=1.2, usearrows=F)

```
# 연계중심성(Betweenness Centrality)
###Edge의 중심성 (임의의 노드간 거리에서 edge를 거치는 빈도)
# 아이겐벡터(Eigen vector Centrality)
###중심성 높은 노드들간 연계
```{r}
options(warn = FALSE)
#Indicate only connectivity.
tele_mat2[tele_mat2 > 0] <- 1
#tele_mat2

#Betweenness Centrality
betweenness(tele_mat2)
#Get Index of Maximum betweenness Centrality
#rownames(tele_mat2)[which.max(betweenness(tele_mat2))]

#EigenVector Centrality

#round(evcent(tele_mat2) , 3)
#round(evcent(t(tele_mat2)) , 3)

#Manual Eigenvector Calclulations
#abs(eigen(tele_mat2)$vectors[,1])
#rownames(tele_mat2)[which.max(abs(eigen(tele_mat2)$vectors[,1]))]

```

```{r}
#변환

email_mat2 <- matrix(0, 12, 12)
email_mat2[email_mat[,1:2]] <- 1
rownames(email_mat2) <- email.name
colnames(email_mat2) <- email.name

```

```{r}

#email_mat2
#getwd()

```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save?the notebook, an HTML f?le containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).
